---
title: "prediction"
author: "Felipe Frazatto"
date: "10/5/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}

library(dplyr)
library(tidyr)

```


# Reading Database file

```{r}

pF <- readRDS("./predFile.rds")

```


# Define Side Functions

## CleanText

Used to remove punctuation, numbers, elongated words and multiple blank spaces.

```{r}

cleanText <- function(text) {
     
     newText <- text %>% 
          tolower() %>% 
          gsub(pattern = "[[:punct:]]", replacement =  " ") %>% 
          gsub(pattern = "[0-9]", replacement = " ") %>% 
          gsub(pattern = "\\b(?=\\w*(\\w){3}\\1)\\w+\\b", replacement = " ", perl = TRUE) %>%
          gsub(pattern = " +", replacement = " ")
          
     return(newText)
     
}

```

## Prep

Prepares input string for the prediction. Cleans the string and separate it 
by word.

```{r}

prep <- function(input.Str){
     
     # Initialize vector. Used for the string splititng into a word data frame
     words <- c()
     
     # Cleans String
     cleanStr <- cleanText(input.Str)
     
     # Counts number of words
     wordCount <- length(strsplit(cleanStr, " ")[[1]])
     
     # Populate words vector
     for(i in 1:wordCount){
          
          # Builds names following the format: "w" + i, where i = 1, 2, ..., n
          words <- c(words, 
                     paste("w", i, sep = ""))
          
     }
     
     
     # Separates the cleaned input string into a word data frame
     strEval <- separate(data.frame(str = cleanStr), 
                         str, 
                         sep = " ", 
                         into = words)
     
     return(strEval)
     
}

```


## ngramFilter


Filters the training data frame with respect to a defined ngram and then to its
words.

```{r}

ngramFilter <- function(strEval, ngram){
     
     # Filters initial pF dataframe by the wanted ngram
     subSet <- filter(pF, n == ngram) 
     
     # Calculates the input string size
     inputSize <- dim(strEval)[2]
     
     # Sequential filtering, word by word
     for(j in 1:(ngram - 1)){
          
          subSet <- subSet %>% 
               filter(.[,j] == strEval[, inputSize - ngram + j + 1])
          
     }
     
     # Calculates the probability for each sequence to occur
     subSet$prob <- subSet$freq/sum(subSet$freq)   
     
     return(subSet)
     
}

```

# Prediction

## pred

The predict function. Prepares the input, filters the data set to a more concise
one, and get the probability of get a sequency.

Important to notice, if the ngramFilter function fail to find any thing like the
input sequence it will return a 0 by 0 data frame, in this case the algorithm 
will sugest a period ".".

```{r}

pred <- function(input.Str){

     # Prepares input string
     strEval <- prep(input.Str)
     
     # Load inital data set
     subSet <- pF
     
     # Finds the longest ngram in the data set
     ngramLim <- max(subSet$n)
     
     # Looks for a possible output prediction. Begins from the longest ngram 
     # available to the shortest one (bigram). The loop will quit as soon as 
     # a valid sequence is found.
     for(i in ngramLim:2){
          
          subSet <- ngramFilter(strEval, ngram = i)
          
          if(dim(subSet)[1] != 0)
          {
               break
          }
     
     }
     
     # If a sequence is found, suggest the word with highest probability (first
     # printed word) and picks randomly a second word, following the Mass
     # Probability Distribution calculated for the sequence.
     if(dim(subSet)[1] != 0){
         
       samples <- 0 
       randProb <- c()
       
       if(dim(subSet)[1] > 1){
        
          randProb <- sample(subSet[,max(subSet$n)], 2, prob = subSet$prob)
       
       }
       
       else{
         randProb <- c(".","")
       }
          
       maxProb <- subSet[which.max(subSet$prob), max(subSet$n)]
          
          
          output <- c(maxProb, randProb)
          
     }
     
     # If no valid sequence is found, suggests a period.
     else{output <- "."}
     
     return(output)

}

```

# Exemples

```{r}

pred("There")

```

```{r}

pred("There is")

```

```{r}

pred("There is a")

```


```{r}

pred("There is a house")

```

```{r}

pred("There is a house in New")

```

# Conlusion


The algorithm is capable to predict some sentences, however it does not know the
song The House of the Rising Sun by the Animals...